---
title: "Trends in age continuity"
output:
  rmarkdown::github_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyverse)


dat <- readRDS("./gen/rates.rds")
v_cod_order <- c("Measles", "Maternal", 
                 "Intrapartum", "Preterm","SepsisMening", "Tetanus", "MeningEnceph" ,
                 "HIV", "ARI","LRI",  "TB", "Diarrhoeal", "Malaria", "OtherCMPN",
                "Congenital", "Cardiovascular", "Digestive", "Neoplasms", "OtherNCD",
                "InterpVio","SelfHarm", "Drowning", "RTI", "OtherInj", "NatDis", "CollectVio",
                "COVID19","OtherCOVID19")
dat$COD <- factor(dat$COD, levels = v_cod_order)
```

# CSMF comparison for 5-9y

### India

* GHE has higher CSMFs for LRI, OtherCMPN, and Diarrhoeal. 
* CA CODE has higher OtherNCD, Digestive, Drowning, and OtherInj.

```{r IND5to9frac}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = frac_cacode_05to09y,
         GHE = frac_ghe_05to09y) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "IND") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="CSMF for 5-9y",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

### Sub-Saharan Africa

Smoothed trend of national-level CSMFs for 5-9y in Sub-Saharan Africa.

* GHE tends to have higher HIV
* CA CODE has higher diarrhoeal, malaria, otherncd

```{r SSA5to9frac}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = frac_cacode_05to09y,
         GHE = frac_ghe_05to09y) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Sub-Saharan Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .1, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="CSMF for 5-9y",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```


# Age continuity in cause-specific rates from 1-4y to 5-9y

* Values above zero mean an increase in the cause-specific rate from 1-4y to 5-9y
* Values below zero mean a decrease in the cause-specific rate from 1-4y to 5-9y

### India

* For GHE, the rate of malaria increases from 1-4y to 5-9y. It is the opposite for CA CODE.
* For GHE, the rate of OtherNCD decreases from 1-4y to 5-9y. The rates are similar between age groups for CA CODE.
* For both GHE and CA CODE, the rate of OtherCMPN is lower in 5-9y than 1-4y.
* For CA CODE, the cause-specific rate of TB was lower for 1-4y in the early 2000s, and higher for 5-9y in the late 2010s.

```{r IND1to9}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)/(rate_cacode_05to09y + rate_ghe_01to04y)/2*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)/(rate_ghe_05to09y + rate_ghe_01to04y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "IND") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

### Nigeria

* Similar age continuity for GHE and CA CODE.
* For both CA CODE and GHE, large decrease in malaria from 1-4y to 5-9y.
* For CA CODE, OtherInj were lower in 5-9y than 1-4y. For GHE the rates are similar between the two age groups.
* Big relative percent difference in Natural Disaster rates from 1-4y to 5-9y for CA CODE. Very small real difference (see below).


```{r NGA1to9}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)/(rate_cacode_05to09y + rate_ghe_01to04y)/2*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)/(rate_ghe_05to09y + rate_ghe_01to04y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "NGA") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

```{r NGA1to9abs}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "NGA") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "Rate difference",
        title="Real difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

### Eastern and South-eastern Asia

Smoothed trend of relative percent differences in national-level rates from 1-4y to 5-9y.

* Similar age continuity for GHE and CA CODE.
* For GHE, RTI was higher in 5-9y than 1-4y. For CA CODE it was the opposite.
* Need to look into flat NatDis for GHE.

```{r ESEasia}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)/(rate_cacode_05to09y + rate_ghe_01to04y)/2*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)/(rate_ghe_05to09y + rate_ghe_01to04y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region ==  "Eastern and South-eastern Asia") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 

```

### Sub-Saharan Africa

Smoothed trend of relative percent differences in national-level rates from 1-4y to 5-9y.

* HIV goes down from 1-4y to 5-9y for both GHE and CA CODE, but more for CA CODE.
* Big relative decrease in NatDis from 1-4y to 5-9y for CA CODE. Small real change.

```{r SSA}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)/(rate_cacode_05to09y + rate_ghe_01to04y)/2*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)/(rate_ghe_05to09y + rate_ghe_01to04y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Sub-Saharan Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 

```

### World

Smoothed trend of relative percent differences in world rates from 1-4y to 5-9y.

* Directionality of change between 1-4y and 5-9y is same for all cause-specific rates between CA CODE and GHE
* For GHE, there is less of a difference between age groups for HIV
* For CA CODE, there is less of a difference between age groups for OtherNCD

```{r world}
p <- dat %>%
  filter(AgeGroup %in% c("01to04y", "05to09y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_cacode_05to09y - rate_ghe_01to04y)/(rate_cacode_05to09y + rate_ghe_01to04y)/2*100,
         GHE = (rate_ghe_05to09y - rate_ghe_01to04y)/(rate_ghe_05to09y + rate_ghe_01to04y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(Name == "World") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 1-4y to 5-9y",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 

```

# CSMF comparison for 15-19y

### India

In India, CA CODE has much lower OtherCMPN and higher OtherNCD in 15-19y than GHE.

```{r IND1519frac}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = (frac_cacode_15to19y),
         GHE = (frac_ghe_15to19y)) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "IND") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "CSMF",
        title="CSMF for 15-19y Total",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

* In India, the GHE CSMFs essentially don't change between 15-19y and 20-24y.
* Because CA CODE has a lower CSMF for OtherCMPN in 15-19y (shown above), there is a large jump to the GHE CSMF (and rate) for 20-24y.
* Because CA CODE has a higher CSMF for OtherNCD in 15-19y (shown above), there is a large drop to the GHE CSMF (and rate) for 20-24y.

```{r IND1524frac}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = (frac_ghe_20to24y - frac_cacode_15to19y),
         GHE = (frac_ghe_20to24y - frac_ghe_15to19y)) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Name == "IND") 
p %>%
  ggplot() + 
  geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "CSMF difference",
        title="Difference in CSMF from 15-19y to 20-24y Total",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

### Sub-Saharan Africa

Smoothed trend of national-level CSMFs for 15-19y.

* GHE has higher Maternal
* CA CODE has higher OtherNCD, InterpVio, SelfHarm
* GHE and CA CODE have different trends in the 15-19y CSMF for OtherCMPN

```{r SSA1519ffrac}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Female") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = frac_cacode_15to19y,
         GHE = frac_ghe_15to19y) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Sub-Saharan Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "CSMF",
        title="CSMF for 15-19y Females",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

* Large differences in 15-19y CSMFs for InterpVio and CollectVio for males

```{r SSA1519mfrac}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Male") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = frac_cacode_15to19y,
         GHE = frac_ghe_15to19y) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Sub-Saharan Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "CSMF",
        title="CSMF for 15-19y Males",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

### Western Asia and Northern Africa

* The 15-19y CSMF for CollectVio is much higher in GHE, InterpVio higher in CA CODE.

```{r WANA1519mfrac}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Male") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("frac_cacode","frac_ghe"))
  ) %>% 
  mutate(CACODE = frac_cacode_15to19y,
         GHE = frac_ghe_15to19y) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Western Asia and Northern Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "CSMF",
        title="CSMF for 15-19y Males",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

# Age continuity in cause-specific rates from 15-19y to 20-24y

* Values above zero mean an increase in the cause-specific rate from 15-19y to 20-24y
* Values below zero mean a decrease in the cause-specific rate from 15-19y to 20-24y

### Sub-Saharan Africa

Smoothed trend of relative percent differences in national-level rates from 15-19y to 20-24y.

* For GHE, the rate of Cardiovascular, Digestive, Neoplasms, and SelfHarm increases from 15-19y to 20-24y. It is the opposite for CA CODE.
* For CA CODE, there are large relative increases in cause-specific rates of CollectVio from 15-19y to 20-24y. 
  * This is due to the CA CODE CSMF for CollectVio being much lower in 15-19y than GHE.

```{r SSA1524mrate}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Male") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_ghe_20to24y - rate_cacode_15to19y)/(rate_ghe_20to24y + rate_cacode_15to19y)/2*100,
         GHE = (rate_ghe_20to24y - rate_ghe_15to19y)/(rate_ghe_20to24y + rate_ghe_15to19y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(AdminLevel == "National" & Region == "Sub-Saharan Africa") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 15-19y to 20-24y - Males",
        subtitle = p %>% distinct(Region)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```


### World

Smoothed trend of relative percent differences in world rates from 15-19y to 20-24y.

* For GHE, the rate of Digestive, OtherNCD increases from 15-19y to 20-24y. It is the opposite for CA CODE.
  * For Digestive, the real difference is rather small (see below)
* For CA CODE, there are large relative increases in CollectVio from 15-19y to 20-24y.
  * In this case, the real difference in rates is notable (see below).


```{r World1524rate}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_ghe_20to24y - rate_cacode_15to19y)/(rate_ghe_20to24y + rate_cacode_15to19y)/2*100,
         GHE = (rate_ghe_20to24y - rate_ghe_15to19y)/(rate_ghe_20to24y + rate_ghe_15to19y)/2*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(Name == "World") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "%",
        title="Relative % difference in cause-specific mx from 15-19y to 20-24y - Total",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

```{r World1524abc}
p <- dat %>%
  filter(AgeGroup %in% c("15to19y", "20to24y") & Sex == "Total") %>% 
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  pivot_wider(
    id_cols = all_of(c("AdminLevel","Name","Region","Sex","Year", "COD")),
    names_from = AgeGroup,
    values_from = all_of(c("rate_cacode","rate_ghe"))
  ) %>% 
  mutate(CACODE = (rate_ghe_20to24y - rate_cacode_15to19y)*100,
         GHE = (rate_ghe_20to24y - rate_ghe_15to19y)*100) %>% 
  filter(!(is.na(CACODE) & is.na(GHE))) %>% 
  select(all_of(c("AdminLevel","Name","Region","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  filter(Name == "World") 
p %>%
  ggplot() + 
  #geom_point(aes(x = Year, y = value, col = Source), alpha = .2, shape = 1) +
  geom_smooth(aes(x = Year, y = value, col = Source)) +
  geom_hline(aes(yintercept = 0)) +
  geom_text(data = . %>% distinct(COD, .keep_all = TRUE), 
            aes(label = COD, y=Inf), 
            x = 2010, hjust = .5, vjust = 1.5, size = 3.5) +
  labs(x="", y= "Rate difference",
        title="Real difference in cause-specific mx from 15-19y to 20-24y - Total",
        subtitle = p %>% distinct(Name)) +
  theme_bw() +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  facet_wrap(~COD) +
  theme(strip.text = element_blank(), legend.direction = "horizontal", legend.position = "bottom") 
```

