---
title: "CA CODE and GHE CSMF Comparison"
subtitle: "Scatter plots and trends 2000-2021"
output:
  html_document:
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyverse)
library(flextable)
library(officer)
library(countrycode)
dat <- readRDS("C:/Users/HEilerts/Institute of International Programs Dropbox/Hallie Eilerts-Spinelli/CA-CODE/Comparison-CACODE-GHE/gen/rates.rds")
v_cod_order <- c("Measles", "Maternal", 
                 "Intrapartum", "Preterm","SepsisMening", "Tetanus", "MeningEnceph" ,
                 "HIV", "ARI","LRI",  "TB", "Diarrhoeal", "Malaria", "OtherCMPN",
                "Congenital", "Cardiovascular", "Digestive", "Neoplasms", "OtherNCD",
                "InterpVio","SelfHarm", "Drowning", "RTI", "OtherInj", "NatDis", "CollectVio",
                "COVID19","OtherCOVID19")
dat$COD <- factor(dat$COD, levels = v_cod_order)
dat$CountryName <- countrycode(dat$Name, origin = "iso3c", destination = "country.name")

#unique(subset(dat, SDGregion == "Central and Southern Asia")$CountryName)
#unique(subset(dat, SDGregion == "Western Asia and Northern Africa")$CountryName)

```

# 5-9y

* GHE has higher HIV in sub-Saharan Africa
* CACODE has higher malaria in sub-Saharan Africa
* GHE has higher collective violence in Central and Southern Africa, Western Asia and Northern Africa

```{r frac5to9_all,  fig.height = 10, fig.width=10}
# fig.height = 8, fig.width=10
pnat <- dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National") &
           Year <= 2021) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe)
pnat %>%
  ggplot() + 
  geom_abline() + 
  geom_point(aes(x=CACODE, y = GHE, color = SDGregion), shape = 1, alpha = .8) +
  labs(title = "5-9y 2000-2021") +
  theme_bw() +
  facet_wrap(~COD, ncol = 5) +
  scale_fill_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  theme(legend.direction = "horizontal", legend.position = "bottom",
        aspect.ratio = 1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

```{r frac5to9_world,  fig.height = 10, fig.width=14, eval = FALSE}
# fig.height = 8, fig.width=10
pnat <- dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National" )&
           Year <= 2021)
pworld <- pnat %>%
  mutate(Name = "World",
         SDGregion = "World")
p <- rbind(pworld, pnat)
p <- p %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe)
p %>% filter(Name == "World") %>%
  ggplot() + 
  geom_abline() + 
  geom_point(aes(x=CACODE, y = GHE), shape = 1) +
  labs(title = "5-9y 2000-2021, World") +
  theme_bw() +
  facet_wrap(~COD, ncol = 6) +
  scale_fill_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  theme(legend.direction = "horizontal", legend.position = "bottom",
        aspect.ratio = 1)
```

## CollectVio

Countries with large differences in CollectVio (CSMFs for GHE > 0.50 and CACODE < 0.10), all causes

```{r bar5to9collectVio,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  mutate(ind = ifelse(COD == "CollectVio"  & GHE > 0.5 & CACODE < 0.1, 1, 0)) %>%
  group_by(CountryName) %>%
  mutate(ind = max(ind)) %>%
  filter(ind == 1) %>%  
  filter(!(CountryName %in% c("Iceland","Antigua & Barbuda"))) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "5-9y", y= "CSMF") +
  facet_grid(CountryName~Source)
```


## HIV

Countries largest average difference in HIV

```{r trend5to9hiv,  fig.height = 6, fig.width=10}
dat %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  filter(COD == "HIV" ) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 10) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  ggplot()  +
  geom_point(aes(x=Year,y=value, col = Source), alpha = .8, shape = 1) +
  geom_smooth(aes(x=Year,y=value, col = Source)) +
  labs(title = "5-9y", y= "CSMF") +
  facet_wrap(~CountryName, ncol = 5) +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "")
```

Countries with largest average difference in HIV, all causes

```{r bar5to9hiv,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  mutate(dif = ifelse(COD == "HIV", dif, 0)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 5) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "5-9y", y= "CSMF") +
  facet_grid(CountryName~Source)

```

## Malaria

Countries largest average difference in malaria

```{r trend5to9malaria,  fig.height = 6, fig.width=10}
dat %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  filter(COD == "Malaria" ) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 10) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  ggplot()  +
  geom_point(aes(x=Year,y=value, col = Source), alpha = .8, shape = 1) +
  geom_smooth(aes(x=Year,y=value, col = Source)) +
  labs(title = "5-9y", y= "CSMF") +
  facet_wrap(~CountryName, ncol = 5) +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") +
  coord_cartesian(ylim = c(0,1))
```

Countries with largest average difference in malaria, all causes

* Sometimes GHE has larger OtherCMPN fraction, sometimes measles and HIV

```{r bar5to9mal,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("05to09y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  mutate(dif = ifelse(COD == "Malaria", dif, 0)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 5) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "5-9y", y= "CSMF") +
  facet_grid(CountryName~Source)

```

# 10-14

* GHE has higher HIV in sub-Saharan Africa
* CACODE has higher malaria in sub-Saharan Africa
* Same as 5-9y

```{r frac10to14_all,  fig.height = 10, fig.width=10}
# fig.height = 8, fig.width=10
pnat <- dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("10to14y") & Sex == "Total" & 
           AdminLevel %in% c("National") &
           Year <= 2021) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe)
pnat %>%
  ggplot() + 
  geom_abline() + 
  geom_point(aes(x=CACODE, y = GHE, color = SDGregion), shape = 1, alpha = .7) +
  labs(title = "10-14y 2000-2021") +
  theme_bw() +
  facet_wrap(~COD, ncol = 5) +
  scale_fill_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  theme(legend.direction = "horizontal", legend.position = "bottom",
        aspect.ratio = 1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

## Malaria

Countries with largest average difference in malaria, all causes

```{r bar10to14malaria,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("10to14y") & Sex == "Total" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  # mutate(ind = ifelse(COD == "Malaria" & GHE < 0.1 & CACODE > 0.35, 1, 0)) %>%
  # group_by(CountryName) %>%
  # mutate(ind = max(ind)) %>%
  # filter(ind == 1) %>%  
  mutate(dif = abs(GHE - CACODE)) %>%
  mutate(dif = ifelse(COD == "Malaria", dif, 0)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 5) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "10-14y", y= "CSMF") +
  facet_grid(CountryName~Source)

```


# 15-19 F

* GHE has higher maternal in sub-Saharan Africa
* CACODE has higher OtherCMPN in sub-Saharan Africa
* GHE has higher CollectVio in western and northern Africa

```{r frac15to19F_all,  fig.height = 10, fig.width=10}
# fig.height = 8, fig.width=10
pnat <- dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("15to19y") & Sex == "Female" & 
           AdminLevel %in% c("National") &
           Year <= 2021) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe)
pnat %>%
  ggplot() + 
  geom_abline() + 
  geom_point(aes(x=CACODE, y = GHE, color = SDGregion), shape = 1, alpha = .7) +
  labs(title = "15-19yF 2000-2021") +
  theme_bw() +
  facet_wrap(~COD, ncol = 5) +
  scale_fill_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  theme(legend.direction = "horizontal", legend.position = "bottom",
        aspect.ratio = 1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

## Maternal

Countries largest average difference in Maternal

```{r trend15to19Fmat,  fig.height = 6, fig.width=10}
dat %>%
  filter(AgeGroup %in% c("15to19y") & Sex == "Female" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  filter(COD == "Maternal" ) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 10) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>%
  ggplot()  +
  geom_point(aes(x=Year,y=value, col = Source), alpha = .8, shape = 1) +
  geom_smooth(aes(x=Year,y=value, col = Source)) +
  labs(title = "15-19yF", y= "CSMF") +
  facet_wrap(~CountryName, ncol = 5) +
  scale_color_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "")
```

Countries with largest average difference in Maternal, all causes.

```{r bar15to19Fmat,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19"))) %>%
  filter(AgeGroup %in% c("15to19y") & Sex == "Female" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>%
  mutate(dif = abs(GHE - CACODE)) %>%
  mutate(dif = ifelse(COD == "Maternal", dif, 0)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 5) %>%
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "15-19yF", y= "CSMF") +
  facet_grid(CountryName~Source)

```

# 15-19 M

* GHE has higher RTI and CollectVio
* CACODE has higher InterpVio

```{r frac15to19M_all,  fig.height = 10, fig.width=10}
# fig.height = 8, fig.width=10
pnat <- dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19","Maternal"))) %>%
  filter(AgeGroup %in% c("15to19y") & Sex == "Male" & 
           AdminLevel %in% c("National") &
           Year <= 2021) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe)
pnat %>%
  ggplot() + 
  geom_abline() + 
  geom_point(aes(x=CACODE, y = GHE, color = SDGregion), shape = 1, alpha = .7) +
  labs(title = "15-19yM 2000-2021") +
  theme_bw() +
  facet_wrap(~COD, ncol = 5) +
  scale_fill_manual(values = c("#8C2981FF","#FA7F5EFF"), name = "") + 
  theme(legend.direction = "horizontal", legend.position = "bottom",
        aspect.ratio = 1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

## CollectVio, InterpVio, RTI

Countries with largest average difference in CollectVio, InterpVio, and RTI.

```{r bar15to19Mcolvio,  fig.height = 6, fig.width=9}
dat %>%
  filter(!(COD %in% c("COVID19", "OtherCOVID19","Maternal"))) %>%
  filter(AgeGroup %in% c("15to19y") & Sex == "Male" & 
           AdminLevel %in% c("National")) %>%
  mutate(CACODE = frac_cacode, GHE = frac_ghe) %>% 
  mutate(dif = abs(CACODE-GHE)) %>%
  mutate(dif = ifelse(COD %in% c("CollectVio","InterpVio","RTI"), dif, 0)) %>%
  group_by(CountryName) %>%
  mutate(dif = mean(dif)) %>%
  mutate(maxdif = max(dif)) %>%
  arrange(-maxdif) %>%
  ungroup() %>%
  mutate(groupRank = dense_rank(-maxdif)) %>% 
  filter(groupRank <= 5) %>% 
  select(all_of(c("AgeGroup","AdminLevel","CountryName","SDGregion","Sex","Year", "COD", "CACODE","GHE"))) %>% 
  pivot_longer(
    cols = all_of(c("CACODE","GHE")),
    names_to = "Source"
  ) %>% 
  ggplot()  +
  geom_bar(aes(x=Year,y=value,fill=COD), stat = "identity") +
  labs(title = "15-19yM", y= "CSMF") +
  facet_grid(CountryName~Source)
#View(subset(dat, CountryName == "Qatar" & AgeGroup == "15to19y" & Sex =="Male")[,c("CountryName","Year","COD","frac_ghe","frac_cacode")])
```

